---
title: "Reverse Engineering"
author: "Janko Thyson"
date: "Saturday, October 30, 2014"
output: html_document
---

```{r, message=FALSE, cache=TRUE}
require("shiny")
options(shiny.suppressMissingContextError=TRUE)
```

## Trying to understand the dependency recognition mechanism

### Classes involved

1. For reactive source objects: `ReactiveValues`
2. For observables:

  - `Observable`
  - `Dependents`
  - `Map`

`Dependents` uses instance of `Map` in field `.dependents`

```{r, eval=FALSE}
Dependents <- R6Class(
  'Dependents',
  portable = FALSE,
  class = FALSE,
  public = list(
    .dependents = 'Map',

    initialize = function() {
      .dependents <<- Map$new()
    },
    register = function(depId=NULL, depLabel=NULL) {
      ctx <- .getReactiveEnvironment()$currentContext()
      if (!.dependents$containsKey(ctx$id)) {
        .dependents$set(ctx$id, ctx)
        ctx$onInvalidate(function() {
          .dependents$remove(ctx$id)
        })

        if (!is.null(depId) && nchar(depId) > 0)
          .graphDependsOnId(ctx$id, depId)
        if (!is.null(depLabel))
          .graphDependsOn(ctx$id, depLabel)
      }
    },
    invalidate = function() {
      lapply(
        .dependents$values(),
        function(ctx) {
          ctx$invalidate()
          NULL
        }
      )
    }
  )
)
```

Important part:

```{r, eval=FALSE}
ctx <- .getReactiveEnvironment()$currentContext()
      if (!.dependents$containsKey(ctx$id)) {
        .dependents$set(ctx$id, ctx)
```


Class `Map`

```{r, eval=FALSE}
Map <- R6Class(
  'Map',
  portable = FALSE,
  public = list(
    initialize = function() {
      private$env <- new.env(parent=emptyenv())
    },
    get = function(key) {
      env[[key]]
    },
    set = function(key, value) {
      env[[key]] <- value
      value
    },
    mset = function(...) {
      args <- list(...)
      if (length(args) == 0)
        return()

      arg_names <- names(args)
      if (is.null(arg_names) || any(!nzchar(arg_names)))
        stop("All elements must be named")

      list2env(args, envir = env)
    },
    remove = function(key) {
      if (!self$containsKey(key))
        return(NULL)

      result <- env[[key]]
      rm(list=key, envir=env, inherits=FALSE)
      result
    },
    containsKey = function(key) {
      exists(key, envir=env, inherits=FALSE)
    },
    keys = function() {
      # Sadly, this is much faster than ls(), because it doesn't sort the keys.
      names(as.list(env, all.names=TRUE))
    },
    values = function() {
      as.list(env, all.names=TRUE)
    },
    clear = function() {
      private$env <- new.env(parent=emptyenv())
      invisible(NULL)
    },
    size = function() {
      length(env)
    }
  ),

  private = list(
    env = 'environment'
  )
)
```

Seems like this is where the information about dependents is actually stored:

```
{Map}$private$env
```

### Current findings and guesses

Looking at the tracing changes made in classes `ReactiveValues2` and `Observable2` this is what I currently understand of or are the current guesses about the dependency recognition mechanism:

- *Something* is assigned to ID value '1' and put in a list element in `.dependents$private$env` (`.dependents` is an instance of `Map`). 

What exactly this something is, that I don't fully understand yet. As I set *two* reactive sources yet there only exists *one* entry, I'm suggesting that this *something* is the instance of `ReactiveValues` which contains all reactive source objects that have been created in the same context?

- The instance of class `Map` is the field value of `Dependents$.dependents`

- The instance of `Dependents` is in turn the field value of `Observable2$.dependents`

- However, I still don't understand how when executing a reactive expression (created via `reactive()`) actually determines/recognizes all of its *dependencies* (which seem to be reffered to as *dependents*? The wording kind of confuses me regarding how depends on what ;-)). 

    - How is that environment in `.dependents$private$env` with ID `1` actually processed/used? 
    - Does it have something to do with the `.graphStack` and its related functions (e.g. `.graphDependsOnId`)? 
    - If so, how can I use this information to see what's going on before everything is invalidated again?

- The decision "from cache" vs. "update" seems to be made in `Observable$getValue()` and has clearly something to do with this *invalidation* mechanism which I also do not really understand yet: if `!.invalidated` (which I'm guessing means "no change in source object") `-->` from cache, else run `$.updateValue()`.

### Reactive source objects

### Changes made for reverse engineering 

```
makeReactiveBinding2() --> reactiveValues2() -- .createReactiveValues2() --> ReactiveValues2 --> field `value_hash.
```

Changed `set` part to add computation of checksum value

#### Note
In `ReactiveValues2` part where `value` and `value_hash` are set via `.graphValueChange()`:
```{r, eval=FALSE}
.graphValueChange(sprintf(`%s$%s`, .label, key), value)
.graphValueChange(sprintf(`%s$%s`, .label, key_hash), value_hash)
```

#### Example

```{r, eval=FALSE, cache.vars=TRUE}
options(shiny.suppressMissingContextError=TRUE)
## Object 1 //
x_1 <- 10
# makeReactiveBinding("x_1")
makeReactiveBinding2("x_1")
x_1

if (FALSE) {
  uid <- .computeObjectUid(id = "x_1", where = environment())
  registry <- getOption("shiny")$.registry
  ls(registry)
  registry[[uid]]$checksum
}
x_1 <- 20
# rm(x_1)

## Object 2 //
x_2 <- 100
# makeReactiveBinding("x_2")
makeReactiveBinding2("x_2")
x_2
# rm(x_2)
```

### Reactive conductors / reactive expressions

#### Changes made for reverse engineering

- `reactive3()` --> `exprToFunction2()` --> `Observable2` --> field `.hash`

#### Example 

```{r, cache.vars=TRUE}
options(shiny.suppressMissingContextError=TRUE)
## Object 3 //
x_3 <- shiny:::reactive3((x_1 + x_2) * 2)
x_3()
```

Trying to figure out where the information about dependencies is drawn from:


```{r}
####################################
## Instance of class `Observable2` #
####################################

## --> assigned as global variable in `reactive3()`

ls(o$.dependents, all.names = TRUE)
ls(o$.dependents$.dependents$private$env)
ls(o$.dependents$.dependents$values())
ls(o$.dependents$.dependents$get("1"))

#######################################
## Instance of class `reactiveValues` #
#######################################

inst
## --> assigned as global variable in "get" operation of `makeReactiveBinding()`

inst$.label
inst$.values$value
## --> corresponds to `reactiveValues` instance of reactive source `x_2`

## Context //
## --> assigned as global variable in `Observable2$.updateValue()`
ls(ctx, all.names = TRUE)
ctx$.domain

## Field `dependents` //
ls(inst$.dependents)

## Field `.allValuesDeps` //
ls(inst$.allValuesDeps, all.names = TRUE)
ls(inst$.allValuesDeps$.dependents$private$env)
inst$.allValuesDeps$register

## Field `.namesDeps` //
ls(inst$.namesDeps, all.names = TRUE)
ls(inst$.namesDeps$.dependents$private$env)

## Field value //
inst$.values$value

x_1 <- 100
x_1
x_3()
x_2 <- 100
x_3()
# rm(x_3)
```

-----

## Questions

1. Q: How/where is the value of `key` determined when calling `{instance-ReactiveValues}$get(key)`?

  Be `x` is an instance of class `ReactiveValues`:
  It appears that running `x` (i.e. `print(x)` or `{env}$x`) will automatically be translated to `x$get(key)` (e.g. in `makeReactiveBinding()`: `values$value`). How is the value of `key` determined?

  A: see S3 wrapper class `reactivevalues` and its S3 methods for `$`, `<-` etc.

```{r}
.createReactiveValues <- function(values = NULL, readonly = FALSE) {
  structure(list(impl=values), class='reactivevalues', readonly=readonly)
}
`$.reactivevalues` <- function(x, name) {
  .subset2(x, 'impl')$get(name)
}
x <- .createReactiveValues()
# envir <- new.env()
# envir$x <- x
# envir$x

x <- structure(list(value = "hello world!", hash = "abcd"), class = "Horst")
`$.Horst` <- function(x, name) {
  .subset2(x, 'hash')
}
envir <- new.env()
envir$x <- x
envir$x
```

